---
title: "Critic vs Audience Movie Ratings Analysis"
author: "Alexander Mona"
date: "2025-01-22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Critic vs Audience Movie Ratings Analysis

This repository contains an analysis of movie ratings from Metacritic and Rotten Tomatoes. The goal is to compare critic and audience reviews, explore differences across genres, and identify patterns in popularity for directors and production companies.

---

## Table of Contents
1. [Data Collection](#data-collection)
    - [Load Rotten Tomatoes Data](#load-rotten-tomatoes-data)
    - [Web Scraping Metacritic Data](#web-scraping-metacritic-data)
2. [Data Cleaning and Merging](#data-cleaning-and-merging)
    - [Merge Rotten Tomatoes with Metacritic Data](#merge-rotten-tomatoes-with-metacritic-data)
    - [Filter and Clean Merged Data](#filter-and-clean-merged-data)
3. [Analysis](#analysis)
    - [Q1: Metacritic vs Rotten Tomatoes](#q1-metacritic-vs-rotten-tomatoes)
        - [Correlation](#correlation)
        - [Scatterplot](#scatterplot)
        - [Histogram](#histogram)
    - [Q2: Genre-Based Score Differences](#q2-genre-based-score-differences)
    - [Q3: Audience vs Rotten Tomatoes](#q3-audience-vs-rotten-tomatoes)
        - [Correlation](#correlation-1)
        - [Scatterplot](#scatterplot-1)
    - [Q4: Popular Directors and Production Companies](#q4-popular-directors-and-production-companies)

---

## Data Collection

### Load Rotten Tomatoes Data


```{r}
rotten_tomatoes <- read.csv("C:/Users/Bok/Documents/rottentomatoes.csv", stringsAsFactors = FALSE)
rotten_tomatoes$summary <- NULL
```

### Web Scraping Metacritic Data

```{r}
library(xml2)
library(rvest)
user_agent <- "Mozilla/5.0 (Windows NT 10.0; Win64; x64) AppleWebKit/537.36 (KHTML, like Gecko) Chrome/120.0.0.0 Safari/537.36"
```

# Initialize an empty data frame to store all movies data

```{r}
all_movies_df <- data.frame(stringsAsFactors = FALSE)

for (i in 0:26) {
  print (i)
  curr_page <- paste("https://www.metacritic.com/browse/movie/all/all/2016/metascore/?page=", i, sep="")
  page <- read_html(curr_page, user_agent)
  
  movie_title <- xml_text(xml_find_all(page, "//div[@class ='c-finderProductCard_title']/h3/span[2]"))
  release_date <- xml_text(xml_find_all(page, "//span[@class ='u-text-uppercase']"))
  meta_score <- as.integer(xml_text(xml_find_all(page, xpath = "//span[@data-v-e408cafe='']")))
  
  summary <- xml_text(xml_find_all(page, xpath = "//div[@class='c-finderProductCard_description']"))
  
  rv <- sample(3:15, 1)
  Sys.sleep(rv)
  
  # Combine all summaries into a single character vector
  all_summaries <- unlist(summary)
  
  movies_df <- data.frame(movie_title, release_date, meta_score, summary = all_summaries)
  all_movies_df <- rbind(all_movies_df, movies_df)
} 
```
## Data Cleaning and Merging
### Merge Rotten Tomatoes with Metacritic Data

```{r}
merged_movie_reviews <- merge(rotten_tomatoes, all_movies_df, by.x = "movie_title", by.y = "movie_title")
```
### Filter and Clean Merged Data

```{r}
library(lubridate)
library(dplyr)

# Extract the year from the original_release_date column
subset_merged <- merged_movie_reviews %>%
  mutate(year = year(mdy(original_release_date))) %>%
  filter(year == 2016)

# Clean the merged data set
subset_merged$year <- NULL

subset_merged$release_date <- NULL

mean_meta_score <- mean(subset_merged$meta_score, na.rm = TRUE)
mean_tomatometer_rating <- mean(subset_merged$tomatometer_rating, na.rm = TRUE)

# For loop to turn any non-ASCII characters to NA
for (col in names(subset_merged)) {
  if (is.character(subset_merged[[col]])) {
    subset_merged[[col]] <- iconv(subset_merged[[col]], to = "ASCII", sub = "")
  }
}

# Clean the data with gsub, removing unknown characters and numbers
subset_merged$directors <- gsub("\\?\\?", "", subset_merged$director)
subset_merged$actors <- gsub("\\?\\?", "", subset_merged$actors)
subset_merged$actors <- gsub("[^a-zA-Z0-9, ]", "", subset_merged$actors)

combo_score <- rowMeans(subset_merged[c("meta_score", "tomatometer_rating")], na.rm = TRUE)
subset_merged$combo_score <- combo_score

subset_merged$tomatometer_rating <- as.numeric(subset_merged$tomatometer_rating)
subset_merged$meta_score <- as.numeric(subset_merged$meta_score)
na.omit(subset_merged)

# Remove the movie still with NA values
subset_merged <- subset_merged[subset_merged$movie_title != "Yosemite", ]
subset_merged <- subset_merged[subset_merged$movie_title != "Equals", ]

# Structure of dataframe
str(subset_merged)

# Write CSV for merged dataframe subset
write.csv(subset_merged, "Subset_Merged.csv")
```

## Analysis
## Q1: Metacritic vs Rotten Tomatoes
### Correlation 
```{r}
cor.test(subset_merged$meta_score, subset_merged$tomatometer_rating)
```
### Scatterplot

```{r}
plot(subset_merged$tomatometer_rating, subset_merged$meta_score,
     xlab = "Tomatometer Rating", ylab = "Metascore",
     main = "Tomatometer vs Metascore Scatterplot",
     col = "red")
```
### Histogram 

```{r}
library(ggplot2)
ggplot(subset_merged, aes(x = tomatometer_rating, fill = "Tomatometer")) +
  geom_histogram(binwidth = 5, color = "black", alpha = 0.7) +
  geom_histogram(aes(x = meta_score, fill = "Metascore"), binwidth = 5, color = "black", alpha = 0.7) +
  labs(title = "Comparison of Tomatometer and Metascore Rating Distribution",
       x = "Rating",
       y = "Frequency") +
  scale_fill_manual(values = c("Tomatometer" = "red", "Metascore" = "blue"))
```

## Q2: Genre-Based Score Differences

```{r}
genre_ratings <- subset_merged %>%
  group_by(genres) %>%
  summarise(avg_rotten_rating = mean(tomatometer_rating, na.rm = TRUE),
            avg_metacritic_rating = mean(meta_score, na.rm = TRUE))

ggplot(genre_ratings, aes(x = genres)) +
  geom_bar(aes(y = avg_rotten_rating, fill = "Rotten Tomatoes"), stat = "identity", position = "dodge", alpha = 0.8) +
  geom_bar(aes(y = avg_metacritic_rating, fill = "Metacritic"), stat = "identity", position = "dodge", alpha = 0.8) +
  labs(title = "Average Ratings by Genre on Rotten Tomatoes vs. Metacritic",
       x = "Genre",
       y = "Average Rating") +
  scale_fill_manual(values = c("Rotten Tomatoes" = "red", "Metacritic" = "blue"), name = "Rating Source") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## Q3: Audience vs Rotten Tomatoes
### Correlation

```{r}
cor.test(subset_merged$audience_rating, subset_merged$tomatometer_rating)
```
### Scatterplot

```{r}
plot(subset_merged$audience_rating, subset_merged$tomatometer_rating,
     xlab = "Audience Rating", ylab = "Tomatometer Rating",
     main = "Audience Ratings vs Tomatometer Scatterplot",
     col = "purple")
```

## Q4: Popular Directors and Production Companies

```{r}
subset_merged$rating_diff <- subset_merged$audience_rating - subset_merged$combo_score

audience_director_ranking <- aggregate(subset_merged$rating_diff, by = list(directors = subset_merged$directors), FUN = mean, na.rm = TRUE)
audience_company_ranking <- aggregate(subset_merged$rating_diff, by = list(production_company = subset_merged$production_company), FUN = mean, na.rm = TRUE)

# Display Top 10 Directors and Companies
top10_audience_directors <- head(audience_director_ranking[order(audience_director_ranking$x, decreasing = TRUE), ], 10)
top10_audience_companies <- head(audience_company_ranking[order(audience_company_ranking$x, decreasing = TRUE), ], 10)
```